---
title: "Discussion_13"
author: "MGinorio"
date: "11/20/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### To Do

Using R, provide the solution for any exercise in either Chapter 4 or Chapter 7 of the calculus textbook.  If you are unsure of your solution, post your concerns.

### Section 4.2: Exercise 9

A 24 ft. ladder is leaning against a house while the base is pulled away at a constant rate of 1 ft/s.

At what rate is the top of the ladder sliding down the side of the house when the base is:

* A: 1 foot from the house?
* B: 10 feet from the house?
* C: 23 feet from the house?
* D: 24 feet from the house?

### Goal

To find the rate of change of two or more variables that are changing with respect to time.

### How?

take the Derivative Implicitly and every variable gets multiplied by $\frac{d}{dt}$ (derivative with respect to time)

### Link Equation to Data 

Pythagorean theorem

$x^2 + y^2 = 24^2$

### Derivative

Pulled Away =  Positive 
$\frac{d}{dt}$ = 1 ft/s


### Derivative with respect to time

$\frac{d}{dt}(x^2) + \frac{d}{dt}(y^2) = \frac{d}{dt}(24^2)$

$2x \frac{dx}{dt} + 2y \frac{dy}{dt} = 0$

$x \frac{dx}{dt} + y \frac{dy}{dt} = 0$

$\frac{dy}{dt} = -\frac{x}{y} \frac{dx}{dt}$

### Solve Pythagorean theorem 

getting the value of a

```{r}
a = sqrt(24^2 - 1^2)
a
```

```{r}
db_dt <- 1
```

### Finding the Derivative of rate of change 

#### A: 1 foot from the house?

```{r}
da_dt <- -1/a
da_dt

```

Result: `r da_dt` ft/s

#### B: 10 feet from the house?

```{r}
b <- 10
a_2 <- sqrt(24^2 - b^2)


da_dt_2 <- (-b * db_dt) / a_2
```

Result: `r da_dt_2` ft/s

#### C: 23 feet from the house?

```{r}
c <- 23
a_3 <- sqrt(24^2 - c^2)
da_dt_3 <- (-c * db_dt) / a_3
```

Result: `r da_dt_3` ft/s

#### D: 24 feet from the house?

```{r}
d <- 24
a_4 <- sqrt(24^2 - d^2)
da_dt_4 <- (-c * db_dt) / a_4
```

Result: `r da_dt_4` ft/s