---
title: 'Assignment #9'
author: "MGinorio"
date: "10/24/2021"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem 1

The price of one share of stock in the Pilsdorff Beer Company (see Exercise 8.2.12) is given by $Y_n$ on the $n$th day of the year. Finn observes that the differences $X_n = Y_{n+1} - Y_n$ appear to be independent random variables with a common distribution having mean $\mu = 0$ and variance $\sigma^2 = 1/4$. If $Y_1 = 100$, estimate the probability that $Y_{365}$ is

## A) $\geq$ 100

$$
P(Y_{365}-Y_1 \geq 0)
$$
```{r}
# Data
y_1 <- 100
mu <- 0
var <- 1/4
sd <- sqrt(var)
y_365 <- 100
x_a <- (y_365 - y_1) / sqrt(365-1)

a_result <- pnorm(x_a, mean = mu, sd = sd, lower.tail=F)
a_result
```


## B) $\geq$ 110
$$
P(Y_{365}-Y_1 \geq 10)
$$

```{r warning=F,message=F}
y_365 <- 110
x_b <- (y_365 - y_1) / sqrt(365-1)
b_result <- pnorm(x_b, mean = mu, sd = sd, lower.tail=F)
b_result
```

## C) $\geq$ 120

$$
P(Y_{365}-Y_1 \geq 20)
$$

```{r warning=F,message=F}
y_365 <- 120
x_c <- (y_365 - y_1) / sqrt(365-1)
c_result <- pnorm(x_c, mean = mu, sd = sd, lower.tail=F)
c_result
```

## Problem 2

Calculate the expected value and variance of the binomial distribution using the moment generating function.

### Binomial Random Variable

The mean and the variance of a random variable X with a binomial probability distribution can be difficult to calculate directly. Although it can be clear what needs to be done in using the definition of the expected value of X and X2, the actual execution of these steps is a tricky juggling of algebra and summations. An alternate way to determine the mean and variance of a binomial distribution is to use the moment generating function for X.

Start with the random variable X and describe the probability distribution more specifically. 

Perform n independent Bernoulli trials, each of which has probability of success p and probability of failure 1 - p. 

Thus the probability mass function is
$$
f(x) = C(n,x) p^x(1-p)^{n-x}
$$
Here the term $C(n , x)$ denotes the number of combinations of n elements taken $x$ at a time, and x can take the values 0, 1, 2, 3, . . ., n.

### MGF - Moment Generating FUnction

Use this probability mass function to obtain the moment generating function of X:
$$
M(t) = E(e^{tX}) = \sum_{x \epsilon S}^{} e^{tx}f(x)
$$
Let X ~ Binom(n,p) then,

$$
M(t) = \sum_{x = 0} {n \choose x}(pe^t)^k (1-p)^{n-k}
$$

$$
M(t) = [pe^t + (1-p)]^n
$$

$$
M'x(t) = npe^t(pe^t +1 - p)^{n-1}
$$
where np = u
and t = 0

$$
M'x(0) = u'_1 = np = E(X)
$$
Second derivative to obtain $E(X^2)$

$$
M''x(t) = n(n-1)(pe^t)^2(pe^t +1 - p)^{n-2} + npe^t(pe^t + 1 - p)^{n-1}
$$
$$
M''x(0) = u'_2 = n(n-1)p^2 + np = E(X^2)
$$

### Variance

$$
Var(X) = u'_2 -(u'_1)^2 =  n(n-1)p^2 + np - n^2p^2
$$
Thus,
$$
Var(X) = np((n-1)p +1 - np) = np(1 - p)
$$
