---
title: "Discussion_10"
author: "MGinorio"
date: "10/30/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Example 11.6 pg. 410

In the Dark Ages, Harvard, Dartmouth, and Yale admitted only
male students. Assume that, at that time, 80 percent of the sons of Harvard men
went to Harvard and the rest went to Yale, 40 percent of the sons of Yale men went
to Yale, and the rest split evenly between Harvard and Dartmouth; and of the sons
of Dartmouth men, 70 percent went to Dartmouth, 20 percent to Harvard, and
10 percent to Yale. We form a Markov chain with transition matrix.

$$
P = 
\begin{bmatrix}
    0.8 & 0 & 0.2 \\
    0.2 & 0.70 & 0.10 \\
    0.3 & 0 & 0.40 \\
    
\end{bmatrix}
$$

## Step 1
Create vector College_Admis containing the percentages of college admission for each graduate.

```{r}
College_Zone <- c("Harv", "Dartm", "Yale")
College_Zone

```


## Step 2

Create Matrix with values from probability.
convert vector into matrix and add to column and row name 

```{r}
Zone_Transition = matrix(c(0.8,0,0.2,0.20,0.70,0.10,0.30,0.30,0.40), 
                         nrow = 3,
                         byrow = TRUE,
                         dimnames = list(College_Zone, College_Zone))
Zone_Transition

```

## Step 3 

Install markovchain package
 
```{r}
library(markovchain)
```

## Step 4

Create a markov Chain object 
state space = to vector in Step 1 and 
Transition matrix from step 2

```{r}
mcZone <- new("markovchain", states = College_Zone,
               byrow = TRUE,
               transitionMatrix = Zone_Transition,
               name = "College_Movement")
mcZone
class(mcZone)
```


## Step 5
For Example 11.6, find the probability that the grandson of a man from Harvard went to Harvard

```{r}
mcZone^2
```

### Answer

The probability that the grandson of a man from Hardvard went to Harvard is **70%**


## Step 7

Determine the stationary state of 

```{r}
steadyStates(mcZone)
```

## Step 8

Display the Markov chain and the transition probabilities.

```{r}
layout <- Zone_Transition
plot(mcZone, node.size = 10, layout = layout)
```

Since we DO NOT have an absorbing Markov chain, we DO NOT calculate the expected time until absorption.
