---
title: "Ginorio_Assignement_2"
author: "MGinorio"
date: "9/5/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem Set 1

1. Show that $A^TA \neq AA^T$ in general. (Proof and demonstration.)

2. For a special type of square matrix $A$, we get $A^TA = AA^T$. Under what conditions
could this be true? (Hint: The Identity matrix I is an example of such a matrix).

## Part 1

Proof of $AA^T \neq A^TA$ in general.

### Proof

Now to prove $A^TA \neq AA^T$ in general for a square matrix.

Given a matrix

$$
A = 
\begin{bmatrix}
    a & b \\
    c & d \\
\end{bmatrix}
$$
With a transpose as
$$
A^T = 
\begin{bmatrix}
    a & c \\
    b & d \\
\end{bmatrix}
$$

Matrix multiply $AA^T$

$$
AA^T = 
\begin{bmatrix}
    a^2+b^2 & ac+bd \\
    ac+bd & c^2+d^2 \\
\end{bmatrix}
$$

Matrix multiply $A^TA$

$$
A^TA = 
\begin{bmatrix}
    a^2+c^2 & ab+cd \\
    ab+cd & b^2+d^2 \\
\end{bmatrix}
$$

Thus

$$
\begin{bmatrix}
    a^2+b^2 & ac+bd \\
    ac+bd & c^2+d^2 \\
\end{bmatrix}
\neq
\begin{bmatrix}
    a^2+c^2 & ab+cd \\
    ab+cd & b^2+d^2 \\
\end{bmatrix}
$$

### Demonstration

$$
A = 
\begin{bmatrix}
    1 & 3 \\
    2 & 4 \\
\end{bmatrix}
$$
With a transpose as
$$
A^T = 
\begin{bmatrix}
    1 & 2 \\
    3 & 4 \\
\end{bmatrix}
$$

Matrix multiply $AA^T$

$$
AA^T = 
\begin{bmatrix}
    10 & 14 \\
    14 & 20 \\
\end{bmatrix}
$$

Matrix multiply $A^TA$

$$
A^TA = 
\begin{bmatrix}
    5 & 11 \\
    11 & 25 \\
\end{bmatrix}
$$

Thus

$$
\begin{bmatrix}
    10 & 14 \\
    14 & 20 \\
\end{bmatrix}
\neq
\begin{bmatrix}
    5 & 11 \\
    11 & 25 \\
\end{bmatrix}
$$

## Part 2

Under what conditions is $AA^T = A^TA$?

Answer: When matrix $A$ is **symmetric**. Thus $A = A^T$. And then, with substitution $AA = AA$.

$$
A = 
\begin{bmatrix}
    1 & 2 \\
    2 & 4 \\
\end{bmatrix}
$$
With a transpose as
$$
A^T = 
\begin{bmatrix}
    1 & 2 \\
    2 & 4 \\
\end{bmatrix}
$$

Matrix multiply $AA^T$

$$
AA^T = 
\begin{bmatrix}
    5 & 10 \\
    10 & 20 \\
\end{bmatrix}
$$

Matrix multiply $A^TA$

$$
A^TA = 
\begin{bmatrix}
    5 & 10 \\
    10 & 20 \\
\end{bmatrix}
$$

Thus

$$
\begin{bmatrix}
    5 & 10 \\
    10 & 20 \\
\end{bmatrix}
=
\begin{bmatrix}
    5 & 10 \\
    10 & 20 \\
\end{bmatrix}
$$
# Problem Set 2

Write an R function to factorize a square matrix $A$ into LU or LDU, whichever you prefer. Please submit your response in an R Markdown document.

You don’t have to worry about permuting rows of $A$ and you can assume that $A$ is less than $5 \times 5$, if you need to hard-code any variables in your code.

## Solution

LU Decomposition
Gaussian Elimination
L = Lower Trialgular Matrix
U = Upper Triangular Matrix

### Conditions

* A must be able to be reduced to row-echelon form
* L and U are not unique
* Build L using the opposites multiplies 

```{r}
library(matrixcalc)

A <- matrix( c ( 1, 2, 2, 1 ), nrow=2, byrow=TRUE)
luA <- lu.decomposition( A )
L <- luA$L
U <- luA$U

print( L )

print( U )
print( L %*% U )
print( A )

```
```{r}
B <- matrix( c( 2, -1, -2, -4, 6, 3, -4, -2, 8 ), nrow=3, byrow=TRUE )
luB <- lu.decomposition( B )
L <- luB$L
U <- luB$U
print( L )
print( U )
print( L %*% U )
print( B )
```



