---
title: "Ginorio_Assignment_5"
author: "MGinorio"
date: "9/26/2021"
output: pdf_document
---

 Problem 2

(Binomial). The probability of your organization receiving a Joint Commission inspection in any given month is .05. What is the probability that, after 24 months, you received exactly 2 inspections? What is the probability that, after 24 months, you received 2 or more inspections? What is the probability that your received fewer than 2 inspections? What is the expected number of inspections you should have received? What is the standard deviation?

### Solution

$$
(Px) = {n \choose x} \times p^2 \times (1-p)^{n-x}
$$
X = 2, p = .05, n = 24

$$
(Px=2) = {24 \choose 2} \times .05^2 \times (.95)^{22}
$$

```{r}
two_insp <- choose(24, 2) * .05^2 * .95^22
```

Exactly two inspections `r two_insp` (or `r two_insp * 100`%).

```{r}
# For exactly 2
dbinom(2, size=24, prob=0.05) 
```

Confirming above with an R function

```{r}
less_than_two <- pbinom(1, size=24, prob=0.05) 
two_or_more <- 1 - less_than_two
```

Two or more inspections `r two_or_more` (or `r two_or_more * 100`%) and less than two inspections `r less_than_two` (or `r less_than_two * 100`%).

```{r}
n <- 24
prob <- .05
x <- 0:24
p <- dbinom(x,n,prob)
(p * 100)
```

The result with the highest probability (or likelihood) is 1.

```{r}
# Compute mean
mu <- sum(x*p)
# Compute variance
sigma.sq=sum(x^2*p)-mu^2
# Compute standarad deviation
sigma=sqrt(sigma.sq)
```

Mean (or expected value) is `r mu`.

Variance is `r sigma.sq`.

Standard deviation is `r sigma`.