---
title: "assmt11"
author: "MGinorio"
date: "11/7/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Regression Analysis

Using the "cars" dataset in R, build a linear model for stopping distance
as a function of speed and replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and residual analysis.)

```{r packages, message=FALSE, warning=FALSE}

library(tidyverse)
library(moderndive)
library(skimr)

```

Let’s consider a simple example of how the speed of a car affects its stopping distance, that is, how far it travels before it comes to a stop. To examine this relationship, we will use the 'cars' dataset.

```{r}

cars_regression <- cars %>% 
  select(speed, dist)

cars_regression %>% skim()
  

```

### Question

How the speed of a car affects its stopping distance, that is, how far it travels before it comes to a stop?

### EDA

***



#### Variables

Exploratory analysis of variables 

$y:$ Dependent Variable - $dist$ 

$\vec{x}$ Independent Variable - $speed$


Investigate correlation between this variables

* distance
* speed

```{r eda vars, echo=FALSE}

#### distance

pl1 <- ggplot(cars_regression, aes(x = dist)) +
  geom_histogram(binwidth = 30) +
  labs(x = "distance", y = "count") +
  ggtitle("Distance Distribution") +
  theme(legend.position = "none")


#### speed

pl2 <- ggplot(cars_regression, aes(x = speed)) +
  geom_histogram(binwidth = 5) +
  labs(x = "speed", y = "count") +
  ggtitle("Speed Distribution") +
  theme(legend.position = "none")


library(patchwork)

pl1 + pl2 + plot_layout(ncol = 2)

```

#### Visualize Data

Initial scatterplot of the stopping distance as a function of speed indicates the stopping distance tends to increase as the speed increases, as is expected. 

The plot does show the relationship is likely linear.


```{r vis data, echo=FALSE}

pl4 <- ggplot(data = cars_regression, aes(x = speed, y = dist)) + 
  geom_point() +
  labs(x = "Speed (in miles per hour)", y = "Stopping Distance (in feet)",
       title = "Stopping Distance as a function of speed") + 
    geom_smooth(method = "lm", se = F)

pl4
```

#### Create Model

The output of the model indicates a linear function as:

$$
stopping\ distance = -17.579 + (3.932 * speed)
$$

A y-intercept of -17.579 does seem peculiar. 

Based on the linear model, this would indicate that a car at speeds 0 or closer to 0 (negative) it would stop in less than 0 feet, which is accurate since a car that is not in movement thus it does not need to stop. 

The slope of 3.932 based on the speed.



```{r cars model, echo=FALSE}

library(knitr)

#cars_model ~ speed

speed_model <- lm(dist ~ speed, data = cars_regression)

get_regression_table(speed_model, conf.level = 0.99) %>% 
  kable()

```

**Coefficient**

also seen under estimate - This portion of the output shows the estimated coefficient values

**Std. Error**

For a good model, we typically would like to see a standard
error that is at least five to ten times smaller than the corresponding coefficient

For Example:

the SD error for $speed$ is 9.45 times smaller then the coefficient value.(3.932/0.416)

**P-value**

shows the probability that the corresponding coefficient is not relevant in the model. This value is also known as the significance or p-value of the coefficient

The probability that the intercept is not relevant is 0.012.


```{r speed model points, echo=FALSE}

speed_model_points <- get_regression_points(speed_model) 

speed_model_points

```

**Residuals**

The residuals are the differences between the actual measured values and
the corresponding values on the fitted regression line.

Residual values are normally distributed around
a mean of zero in this case we see that even though the values are not exactly zero we can still expect a normal distribution.

That is, a good model’s residuals should be roughly balanced around and not
too far away from the mean of zero.

**RSquared & Residual Standard Error RSE**

These final few lines in the output provide some statistical information
about the quality of the regression model’s fit to the data

```{r}
summary(speed_model)$r.squared

summary(speed_model)$sigma
```


**Minimum Maximum** 

minimum and maximum values of roughly the
same magnitude, and first and third quartile values of roughly the same
magnitude.

```{r}
summary(speed_model_points$residual)
```


**Residual Visual Relationship**

Distribution of Residuals 
Investigate potential relationships between the residuals and all explanatory/predictor variables

**Residual Vs Fitter**

we may be able to construct a model that produces tighter residual
values and better predictions.

Residual values greater than zero mean that the regression model predicted a value that was too small compared to the actual measured value, and negative values indicate that the regression model predicted a value that was too large


**QQ Plot**

If the residuals were normally distributed, we would expect the points
plotted in this figure to follow a straight line. Which in this case we do see a straight line forming.

This test could confirm that the speed as a predictor in the model may be sufficient to explain the data.


```{r residuals, echo=FALSE}

# Code to visualize distribution of residuals:

pl5 <- ggplot(speed_model_points, aes(x = residual)) +
  geom_histogram(bins = 20) +
  labs(x = "speed Residual", y = "Count")

# Code to visualize partial residual plot over speed:

pl6 <- ggplot(speed_model_points, aes(x = speed, y = residual)) +
  geom_point() +
  labs(x = "speed", y = "Residual")


# Code to visualize partial residual plot over dist:

pl7 <- ggplot(speed_model_points, aes(x = dist, y = residual)) +
  geom_point() +
  labs(x = "Distance", y = "Residual")


library(patchwork)

pl6 + pl5 + pl7 + plot_layout(ncol = 2)

```

```{r}
plot(speed_model)
plot(speed_model_points)

```


### Predictions

We do this so that we can specify that 8 is a value of speed, so that predict knows how to use it with the model stored in speed_model

```{r}
predict(speed_model, newdata = data.frame(speed = c(8, 21, 50)))
```

$stopping\ distance = -17.579 + (3.932 * speed)$

$stopping\ distance = -17.579 + (3.932 * 8)$

$stopping\ distance = -17.579 + (3.932 * 21)$

$stopping\ distance = -17.579 + (3.932 * 50)$ ** note 50 is out of range

### Conclusion

Overall, the car speed would appear to be a good predictor of stopping distance. The linear regression model does contain some flaws, particularly in the intercept value and the predictions at higher speeds

