---
title: ""
author: "MGinorio"
date: "8/27/2021"
output: 
  html_document: default
  pdf_document: default
---
## HW_1

One of the most useful applications for linear algebra in data science is image manipulation.  We often need to compress, expand, warp, skew, etc. images.  To do so, we left multiply a transformation matrix by each of the point vectors.  

For this assignment, build the first letters for both your first and last name using point plots in R.  For example, the following code builds an H.

## Practice {.tabset .tabset-pills}
### Example

```{r}
x=c(rep(0,500),seq(0,1,length.out=1000), rep(1,500))

y=c(seq(-1,1,length.out=500),rep(0,1000), seq(-1,1,length.out=500))

z=rbind(x,y)

plot(y~x, xlim=c(-3,3), ylim=c(-3,3))
```

### Task


write R code that will left multiply (%>%) a square matrix (x) against each of the vectors of points (y). Initially, that square matrix will be the Identity matrix.  

Use a loop that changes the transformation matrix incrementally to demonstrate 1) shear, 2) scaling, 3)  rotation , and 4) projection in animated fashion.  

Hint:  Use x11() to open a new plotting window in R.

Upload your document as a .RMD file.  I will know if your assignment is correct if the animation runs. correctly


```{r}
x=c(rep(-1,500), seq(-1,0,length.out=1000), seq(-1,0,length.out=1000),
    rep(0,500), seq(0.25,.75,length.out=500), rep(.25,1000), seq(.75,1.25,length.out=500))

y=c(seq(-1,1,length.out=500), rep(0,1000), rep(1,1000), seq(-1,1,length.out=500), 
    rep(1,500), seq(-1,1,length.out=1000), rep(1,500))

z=rbind(x,y)

# plot for Initial letters

plot(y~x, xlim=c(-3,3), ylim=c(-3,3))
```

```{r}
# create an Identity Matrix
# basic R - use diag prompt to create identity matrix

id_matrix <- diag(2)

# create transformation matrix
# slant letter to the right by multiplying by an x-shear Matrix A
# A = [1  0.2]
#     [0   1 ]

# 1) Shear

Matrix_A <- matrix(c(1,0,0.2,1), nrow = 2, ncol = 2)

# 2) Scaling

# Make slanted matrix Narrower
# Apply x-scale Matrix B
# B = [0.8   0]
#     [0     1]

Matrix_B <- matrix(c(0.8,0,0,10), nrow = 2, ncol = 2)


# 3) Rotation

# Rotate by 30 degrees
# Apply Rotation Matrix
# R = [0.866  -0.5]
#     [0.5   0.866]

Matrix_R <- matrix(c(0.866,0.5,-0.5,0.866),nrow = 2, ncol = 2)

# 4) Projection

Matrix_P <- matrix(c(0,1,0,1), nrow = 2, ncol = 2)

# COmbine all matrices to be able to see animation
# use of lists
# Transform_list

Trans_list <- list(id_matrix, Matrix_A, Matrix_B, Matrix_R, Matrix_P)

# hint
# use x11()

x11(title = "HW1_Animation")

# Create loop

for (idx in 1:5) {
  z <- Trans_list[[idx]] %*% z
  plot(z[2,]~z[1,], xlim = c(-10,10), ylim=c(-10,10))
  
}

# use tydiverse more modern version
```




