---
title: "Linear Regression"
author: "MGinorio"
output:
  html_document: default
  pdf_document: default
---

## Credit Card Balance Data

## {.tabset .tabset-pills}

### Overview 

 **Credit Dataset**

**A simulated data set containing information on ten thousand customers. The aim here is to predict which customers will default on their credit card debt**

**Format**

* A data frame with 10000 observations on the following 4 variables.
* ID Identification
* Income Income in $1,000’s
* Limit Credit limit
* Rating Credit rating
* Cards Number of credit cards
* Age Age in years
* Education Number of years of education
* Gender A factor with levels Male and Female
* Student A factor with levels No and Yes indicating whether the individual was a student
* Married A factor with levels No and Yes indicating whether the individual was married
* Ethnicity A factor with levels African American, Asian, and Caucasian indicating the individual’s ethnicity
* Balance Average credit card balance in $.


### Packages

```{r load-packages, message=FALSE, warning=FALSE}

#Packages used

library(tidyverse)
library(moderndive)
library(skimr)
library(ISLR)

```

### Dataset 
#### {.tabset}
##### Credit

Two Numerical Explanatory Variables

```{r}

library(ISLR)

credit_dis10 <- Credit %>% as_tibble() %>% 
  select(ID, debt = Balance, credit_limit = Limit,
         income = Income, credit_rating = Rating, age = Age)

glimpse(credit_dis10)
```

Take a sample n= 10

```{r}
credit_dis10 %>% sample_n(size = 10)
```

#### {-}

### Statistics



```{r statistics ,message=FALSE}

credit_dis10 %>% select(debt, credit_limit, income) %>% skim()

```

**Note**

**Debt**

Debt Mean = $520.00 USD

Debt Median = $460.00 USD

25% Card Holders have debt = $68.75$ or less

**credit_limit**

Mean = $4736.00

Median = $4622.50

75% card holders had incomes of $57,470$ or less

outcome variable = debt
explanatory variables = credit_limit + income 

**Visual Relationship**

**debt ~ credit **

```{r}
ggplot(data = credit_dis10, mapping = aes(x = debt, y = credit_limit)) + 
  geom_point() +
  labs(x = "Credit limit (in $)", y = "Credit Card debt (in $)",
       title = "Debt and Credit Limit") + 
    geom_smooth(method = "lm", se = F)
```

**debt ~ income** 

```{r}
ggplot(data = credit_dis10, mapping = aes(x = debt, y = income)) + 
  geom_point() +
  labs(x = "Income (in $1000)", y = "Credit Card debt (in $)",
       title = "Debt and Income") + 
    geom_smooth(method = "lm", se = F)
```


**Correlation Coefficient**

```{r}
credit_dis10 %>% 
  select(debt, credit_limit, income) %>% 
  cor()
```

**Note** 

debt ~ credit limit = 0.862

**strong positive linear relationship**

more credit limit =  more debt

***

debt ~ income = 0.464

**weak positive linear relationship**

***

credit_limit ~ income = 0.792

**collinearity**

What to do when collinearity found?

### Regression
#### {.tabset}

**Model Creation**

Regression Model Fit = lm(y ~ x1 + x2, data)

```{r}
debt_model <- lm(debt ~ credit_limit + income, data = credit_dis10 )

# get regression table 

get_regression_table(debt_model)
```

**Note**

***

Intercept = $385.179$

**Regression Plane**

***

credit_limit = $0.264$ (positive)

**credit limit has associated increase credit card debt**

correlation $~~not = ~~$ causation

***

income = $-7.66$

**income has associated decrease of credit card debt**

for every increase in each unit of income($1000)


***

**Matching Results**

new explanatory variables

$x_1$ = credit_rating

$x_2$ = age

```{r}
lm(debt ~ credit_rating + age, data = credit_dis10)
get_regression_table(lm(debt ~ credit_rating + age, data = credit_dis10))

```


