---
title: "Linear Regression-1"
author: "MGinorio"
date: "11/3/2021"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE, warning=FALSE}

library(tidyverse)
library(moderndive)
library(skimr)
library(ISLR)
library(tinytex)

```

## Dataset

**evals dataset**

Researchers at the University of Texas in Austin, Texas (UT Austin) tried  to answer the following research question: what factors explain differences in  instructor teaching evaluation scores? 

To this end, they collected instructor  and course information on 463 courses. A full description of the study can be  found at openintro.org.  

```{r}
glimpse(evals)
```

### Question

Can we explain differences in teaching evaluation score based on various teacher attributes?


### Variables

$y:$ Average teaching $score$ based on students evaluations

$\vec{x}$ Attributes like gender, ethnicity, bty_avg


Investigate correlation between this variables

* ID
* Score
* Age
* Gender
* Rank
* Cls_students

```{r}
evals_disc12 <- evals %>% 
  select(ID, score, gender, age, rank, cls_students, ethnicity)

evals_disc12

```

### EDA

Understanding Each Variable

We will perform an exploratory analysis of the selected variables before any 
formal modeling

```{r}
library(ResourceSelection)

kdepairs(evals_disc12==)
```



```{r eda vars, echo=FALSE}

#### age

pl1 <- ggplot(evals_disc12, aes(x = age)) +
  geom_histogram(binwidth = 5) +
  labs(x = "age", y = "count") +
  ggtitle("Age Distribution") +
  theme(legend.position = "none")


#### score

pl2 <- ggplot(evals_disc12, aes(x = score)) +
  geom_histogram(binwidth = 5) +
  labs(x = "score", y = "count") +
  ggtitle("Score Distribution") +
  theme(legend.position = "none")


#### gender <fct>


pl3 <- ggplot(evals_disc12, aes(x = gender)) +
  geom_bar() +
  labs(x = "gender", y = "count") +
  ggtitle("Gender Distribution") +
  theme(legend.position = "none")


#### cls_students


pl4 <- ggplot(evals_disc12, aes(x = cls_students)) +
  geom_histogram(binwidth = 5) +
  labs(x = "cls_students", y = "count") +
  ggtitle("Students in Class Distribution") +
  theme(legend.position = "none")


### ethnicity <fct>

pl5 <- ggplot(evals_disc12, aes(x = ethnicity)) +
  geom_bar() +
  labs(x = "ethnicity", y = "count") +
  ggtitle("Ethnicity Distribution") +
  theme(legend.position = "none")


### rank <fct>

pl6 <- ggplot(evals_disc12, aes(x = rank)) +
  geom_bar() +
  labs(x = "rank", y = "count") +
  ggtitle("Rank Distribution") +
  theme(legend.position = "none")


library(patchwork)

pl2 + pl1 + pl3 + pl4 + pl5 + pl6 +  plot_layout(ncol = 2)

```

### Statistics

```{r}
evals_disc12 %>% 
  select(score, gender, age, rank, cls_students, ethnicity) %>% skim()
```



**Visual Relationship**

Following the scatterplots indicate the relationship of the independent variables witht the dependent variable (score)

```{r, echo=FALSE}

### age

pl7 <- ggplot(data = evals_disc12, aes(x = age, y = score)) + 
  geom_point() +
  labs(x = "Age", y = "Teaching Score",
       title = "Score vs. Age ") + 
    geom_smooth(method = "lm", formula = y~x)

### gender <fct>

pl8 <- ggplot(data = evals_disc12, aes(x = gender, y = score)) + 
  geom_point() +
  geom_categorical_model() + 
  labs(x = "gender", y = "Teaching Score",
       title = "score and gender")
    

### students in class

pl9 <- ggplot(data = evals_disc12, aes(x = cls_students, y = score)) + 
  geom_point() +
  labs(x = "cls_students", y = "Teaching Score",
       title = "score vs. class students") + 
    geom_smooth(method = "lm", formula = y~x)

### rank <fct>

pl10 <- ggplot(data = evals_disc12, aes(x = rank, y = score)) + 
  geom_point() +
  geom_categorical_model() + 
  labs(x = "rank", y = "Teaching Score",
       title = "score vs. rank")


### ethnicity <fct>

pl11 <- ggplot(data = evals_disc12, aes(x = ethnicity, y = score)) + 
  geom_point() +
  geom_categorical_model() + 
  labs(x = "ethnicity", y = "Teaching Score",
       title = "score vs. ethnicity")


library(patchwork)

pl7 + pl8 + pl9 + pl10 + pl11 + plot_layout(ncol = 2)

```

* Score = 
* Age = 
* Cls_students = 

CATEGORICAL VARS  - 

When using a categorical predictor variable, the intercept corresponds to the mean for the baseline group, while coefficients for the non-baseline groups are offsets from this baseline. Thus in the visualization the baseline for comparison group's median is marked with a solid line, whereas all offset groups' medians are marked with dashed lines

* Gender = 
* Rank = 
* Ethnicity = 

### Dichotomous Var

We can see in the scatter plot above that the amount of students in the classrooms goes from 0-600. In which the vast majority is distributed between 0-200. Having that in mind I will create a Dichotomous variable for num_students in a classroom less than 100 for "Y" and more than 100 to "N"

```{r}
evals_disc12$num_students <- ifelse(evals_disc12$cls_students <= 100, "Y", "N")
evals_disc12$num_students_n <- ifelse(evals_disc12$num_students == "Y", 1,0)
```

```{r}
evals_disc12 %>% 
  ggplot(aes(x = num_students, y = score)) +
  geom_boxplot() + 
  labs(title = "num_students vs. score")
```
num_students boxplot indicates a slight difference in score when teacher has more than 100 students 




### Create Model

**Score Model**

```{r score model, echo=FALSE}

#score_model ~ bty_avg

score_model <- lm(score ~ gender + age + rank + cls_students + ethnicity, data = evals_disc12)

summary(score_model)
 

```

```{r}
plot(score_model)
```


### Transform Variables

**Quadratic Term**

I decided to create quadratic term for class students

```{r}
evals_disc12$cls_students_sq2 <- evals_disc12$cls_students^2
```


**Dichonomous by quatitative**

to create a dichotomous by quatitative variable. I will multiply age by a Dichotomous variable 'num_students'

num_students was created earlier based on the scatterplot

```{r}
evals_disc12$cls_by_age <- evals_disc12$num_students_n * evals_disc12$age

```


### Second Model


**New Variables**

* cls_by_age

* num_students_n

* num_students


```{r}
score_model_t <- lm(score ~ cls_by_age + num_students + cls_students_sq2, data = evals_disc12)

summary(score_model_t)
```

```{r}
plot(score_model_t)
```


## Conclusion

Overall, Adjusted R-squared:  0.009235 wich means the model only accounts for 0.9% of the variability in the data.

the selection of class size (cls_students) given the influence of the outliers.

this model does not meet the level of the baseline model.
