---
title: "Ginorio_Assignment10"
author: "MGinorio"
date: "10/31/2021"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.
A guard agrees to make a series of bets with him. If Smith bets A dollars,
he wins A dollars with probability .4 and loses A dollars with probability .6.
Find the probability that he wins 8 dollars before losing all of his money if

(a) he bets 1 dollar each time (timid strategy).
(b) he bets, each time, as much as possible but not more than necessary to
bring his fortune up to 8 dollars (bold strategy).
(c) Which strategy gives Smith the better chance of getting out of jail

$q_z = \frac{(q/p)^z-1}{(q/p)^M-1}$

Starting with 1 dollar = $z=1$.
$q_z$
$p$  = the probability of a win, and 
$q$ = the probability of a loss.
$M$ = reaches the value M or the value 0


### (a) 

Find the probability that he reaches M = 8 dollars before losing M = 0

He bets 1 dollar each time 
(timid strategy)

```{r timid}

p <- .4
q <- 1 - p
z <- 1
M <- 8

result_timid <- ((q/p)^z - 1) / ((q/p)^M - 1)

result_timid

```

The probability of winning 8 dollars is 0.02030135

### (b) 
he bets, each time, as much as possible 
but not more than necessary
to bring his fortune up to 8 dollars 
(bold strategy).

In order for Smith to bring his fortune to 8 dollars he needs to double down 3 times

$$
1^2 = 2
$$

$$2^2 = 4$$
$$2^3 = 8$$

```{r bold}
result_bold <- p^3

result_bold
```
The probability of winning 8 dollars is 0.064

```{r}
# random walk
# origin
x = 1
y = 0
no_of_points <- 1000

for (i in 1:no_of_points){
  # create random number between 1 to 4
  s = sample(1:4,1)
  if(s==1){ # right x+1
    x = c(x,x[length(x) +1])
    y = c(y,y[length(y)] -1)
  }
  
  if(s==2){ # right x+1
    x = c(x,x[length(x) +2])
    y = c(y,y[length(y)] +2)
  }
  
  if(s==3){ # right x+2
    x = c(x,x[length(x) +4])
    y = c(y,y[length(y)] +4)
  }
  
}

plot(x,y,col="black", type = "l")
```


### (c)

Which strategy gives Smith the better chance of getting out of jail?

Based on Smith arrangement the **bold strategy** is the best strategy it gives Smith 6% chance meanwhile the timid strategy gives Smith 2% chance.
